Nic Steyn

### 2025-03-25 site/index.qmd
update with new smc preprint

My thesis focusses on robust statistical methods for analysing infectious disease dynamics with an emphasis on flexible methodology and [valid uncertainty quantification](https://github.com/nicsteyn2/RobustRtEstimators). My research combines Bayesian inference, [sequential Monte Carlo methods](http://nicsteyn2.github.io/SMCforRt/), and epidemiological modelling, with applications to standard epidemiological data, as well as survey data and [wastewater-based surveillance](https://github.com/nicsteyn2/NZWastewaterModelling).

During my studies, I have also worked as a research assistant in the [*The Global Reference Group for Children Affected by Crisis*](https://www.spi.ox.ac.uk/the-global-reference-group-on-children-affected-by-covid-19#collapse3701021), estimating orphanhood in Brazil during the COVID-19 pandemic (publication coming soon), and have held a variety of tutoring positions (departmental, college, and inter-departmental courses).

>>>>

My thesis focuses on robust statistical methods for analysing infectious disease dynamics with an emphasis on flexible methodology and [valid uncertainty quantification](https://github.com/nicsteyn2/RobustRtEstimators). My research combines Bayesian inference, [sequential Monte Carlo methods](http://nicsteyn2.github.io/SMCforRt/), and epidemiological modelling, with applications to standard epidemiological data, as well as survey data and [wastewater-based surveillance](https://github.com/nicsteyn2/NZWastewaterModelling).

During my studies, I have also worked as a research assistant in the [*The Global Reference Group for Children Affected by Crisis*](https://www.spi.ox.ac.uk/the-global-reference-group-on-children-affected-by-covid-19#collapse3701021), estimating [orphanhood in Brazil](https://doi.org/10.1101/2025.01.31.25321479) during the COVID-19 pandemic, and have held a variety of tutoring positions (departmental, college, and inter-departmental courses).

### 2025-03-25 site/papers.qmd
update with new smc preprint

**Steyn, N.**, & Parag, K. V. Robust uncertainty quantification in popular estimators of the instantaneous reproduction number. *medRxiv preprint*. [https://doi.org/10.1101/2024.10.22.24315918](https://doi.org/10.1101/2024.10.22.24315918)

**Steyn, N.**,  Parag, K. V., & Donnelly, C. A. A primer on inference and prediction with epidemic renewal models using sequential Monte Carlo methods. [Link to PDF](https://github.com/nicsteyn2/SMCforRt/blob/main/workingpaper.pdf)

**Steyn, N.**, Unwin, H. J. T., Ponmattam, J., et al. Regional and national estimates of children affected by all-cause and COVID-19-associated orphanhood and caregiver death in Brazil, by age and family circumstance. *medRxiv preprint*. [https://doi.org/10.1101/2025.01.31.25321479](https://doi.org/10.1101/2025.01.31.25321479)

**Steyn, N.**, Chadeau-Hyam, M., Whitaker, M., et al. Pandemic-risk-related behaviour change in England from June 2020 to March 2022: REACT-1 study among over 2 million people. *medRxiv preprint*. [https://doi.org/10.1101/2025.03.03.25323250](https://doi.org/10.1101/2025.03.03.25323250)

@@

Watson, L. M., Plank, M. J., Armstrong, B. A., Chapman, J. R., Hewitt, J., Morris, H., Orsi, A., Bunce, M., Donnelly, C. A., & **Steyn, N**. Jointly estimating epidemiological dynamics of Covid-19 from case and wastewater data in Aotearoa New Zealand. *Communications Medicine*. [https://doi.org/10.1038/s43856-024-00570-3](https://doi.org/10.1038/s43856-024-00570-3)

@@

Binny, R. N., Priest, P., French, N. P., Parry, M., Lustig, A., Hendy, S. C., Maclaren, O. J., Ridings, K. M., **Steyn, N.**, Vattiato, G., & Plank, M. J. Sensitivity of Reverse Transcription Polymerase Chain Reaction Tests for Severe Acute Respiratory Syndrome Coronavirus 2 Through Time. *The Journal of Infectious Diseases*. [https://doi.org/10.1093/infdis/jiac317](https://doi.org/10.1093/infdis/jiac317)

Elliott, P., Whitaker, M., Tang, D., Eales, O., **Steyn, N.**, Bodinier, B., Wang, H., Elliott, J., Atchison, C., Ashby, D., Barclay, W., Taylor, G., Darzi, A., Cooke, G. S., Ward, H., Donnelly, C. A., Riley, S., & Chadeau-Hyam, M. Design and Implementation of a National SARS-CoV-2 Monitoring Program in England: REACT-1 Study. *American Journal of Public Health*. [https://doi.org/10.2105/AJPH.2023.307230](https://doi.org/10.2105/AJPH.2023.307230)

@@

Elliott, P., Eales, O., **Steyn, N.**, Tang, D., Bodinier, B., Wang, H., Elliott, J., Whitaker, M., Atchison, C., Diggle, P. J., Page, A. J., Trotter, A. J., Ashby, D., Barclay, W., Taylor, G., Ward, H., Darzi, A., Cooke, G. S., Donnelly, C. A., & Chadeau-Hyam, M. Twin peaks: the Omicron SARS-CoV-2 BA. 1 and BA. 2 epidemics in England. *Science*. [https://doi.org/10.1126/science.abq4411](https://doi.org/10.1126/science.abq4411)

Plank, M. J., James, A., Lustig, A., **Steyn, N.**, Binny, R. N., & Hendy, S. C. Potential reduction in transmission of COVID-19 by digital contact tracing systems: a modelling study. *Mathematical Medicine and Biology*. [https://doi.org/10.1093/imammb/dqac002](https://doi.org/10.1093/imammb/dqac002)

**Steyn, N.**, Lustig, A., Hendy, S. C., Binny, R. N., & Plank, M. J. Effect of vaccination, border testing, and quarantine requirements on the risk of COVID-19 in New Zealand: A modelling study. *Infectious Disease Modelling*. [https://doi.org/10.1016/j.idm.2021.12.006](https://doi.org/10.1016/j.idm.2021.12.006)

**Steyn, N.**, Plank, M. J., Binny, R. N., Hendy, S. C., Lustig, A., & Ridings, K. A COVID-19 vaccination model for Aotearoa New Zealand. *Scientific Reports* [https://doi.org/10.1038/s41598-022-06707-5](https://doi.org/10.1038/s41598-022-06707-5)

@@

Hendy, S., **Steyn, N.**, James, A., Plank, M. J., Hannah, K., Binny, R. N., & Lustig, A. Mathematical modelling to inform New Zealand’s COVID-19 response. *Journal of the Royal Society of New Zealand*. [https://doi.org/10.1080/03036758.2021.1876111](https://doi.org/10.1080/03036758.2021.1876111)

James, A., Plank, M. J., Hendy, S., Binny, R., Lustig, A., **Steyn, N.**, Nesdale, A., & Verrall, A. Successful contact tracing systems for COVID-19 rely on effective quarantine and isolation. *PLoS One*. [https://doi.org/10.1371/journal.pone.0252499](https://doi.org/10.1371/journal.pone.0252499)

**Steyn, N.**, Binny, R. N., Hannah, K., Hendy, S., James, A., Lustig, A., Ridings, K., Plank, M. J., & Sporle, A. Māori and Pacific People in New Zealand have higher risk of hospitalisation for COVID-19. *New Zealand Medical Journal*. [Link to PDF](https://nzmj.org.nz/media/pages/journal/vol-134-no-1538/maori-and-pacific-people-in-new-zealand-have-a-higher-risk-of-hospitalisation-for-covid-19/55602a1fe7-1696476650/maori-and-pacific-people-in-new-zealand-have-a-higher-risk-of-hospitalisation-for-covid-19-open-access.pdf)

James, A., Plank, M. J., Binny, R. N., Lustig, A., Hannah, K., Hendy, S. C., & **Steyn, N.** A structured model for COVID-19 spread: modelling age and healthcare inequities. *Mathematical Medicine and Biology*. [https://doi.org/10.1093/imammb/dqab006](https://doi.org/10.1093/imammb/dqab006)

Binny, R. N., Baker, M. G., Hendy, S. C., James, A., Lustig, A., Plank, M. J., Ridings, K. M., & **Steyn, N.** Early intervention is the key to success in COVID-19 control. *Royal Society Open Science*. [https://doi.org/10.1098/rsos.210488](https://doi.org/10.1098/rsos.210488)

**Steyn, N.**, Plank, M. J., James, A., Binny, R. N., Hendy, S. C., & Lustig, A. Managing the risk of a COVID-19 outbreak from border arrivals. *Journal of the Royal Society Interface*. [https://doi.org/10.1098/rsif.2021.0063](https://doi.org/10.1098/rsif.2021.0063)

Plank, M., James, A., & **Steyn, N.** Comment: weekly COVID-19 testing with household quarantine and contact tracing is feasible and would probably end the epidemic[^1]. *Royal Society Open Science*. [https://doi.org/10.1098/rsos.201546](https://doi.org/10.1098/rsos.201546)

@@

James, A., Plank, M. J., Hendy, S., Binny, R. N., Lustig, A., & **Steyn, N.** Model-free estimation of COVID-19 transmission dynamics from a complete outbreak. *PLOS ONE*. [https://doi.org/10.1371/journal.pone.0238800](https://doi.org/10.1371/journal.pone.0238800)

@@

**Steyn, N.**, Binny, R. N., Hannah, K., Hendy, S., James, A., Kukutai, T., Lustig, A., McLeod, M., Plank, M. J., Ridings, K., & Sporle, A. Estimated inequities in COVID-19 infection fatality rates by ethnicity for Aotearoa New Zealand. *New Zealand Medical Journal*. [Link to PDF](https://nzmj.org.nz/media/pages/journal/vol-133-no-1521/estimated-inequities-in-covid-19-infection-fatality-rates-by-ethnicity-for-aotearoa-new-zealand/9bb9847717-1696476126/estimated-inequities-in-covid-19-infection-fatality-rates-by-ethnicity-for-aotearoa-new-zealand-open-access.pdf)

@@

**Steyn, N.**, Plank, M. J., & Hendy, S. C. Modelling to support a future COVID-19 strategy for Aotearoa New Zealand. (2021). [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2017/01/modelling-to-support-a-future-covid-19-strategy-for-aotearoa-new-zealand.pdf)

**Steyn, N.**, Hendy, S., Plank, M., & Binny, R. Technical report: modelling the potential spread of COVID-19 during the August 2021 outbreak. (2021).  [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2017/01/technical-report-28-august.pdf)

James, A., Binny, R. N., Hendy, S. C., Lustig, A., **Steyn, N.**, & Plank, M. J. Economic comparison of the use of Alert Levels 3 and 4 in eliminating the Auckland August outbreak: a cost-effectiveness analysis 21 October 2020. (2020). [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2020/11/Economic-comparison-Auckland-August-outbreak_forTPMwebpage_21Oct.pdf)

Binny, R. N., Hendy, S. C., James, A., Lustig, A., Plank, M. J., & **Steyn, N.** Probability of elimination for COVID-19 in Aotearoa New Zealand. (2020). [https://doi.org/10.1101/2020.08.10.20172361](https://doi.org/10.1101/2020.08.10.20172361)

Binny, R. N., Lustig, A., Brower, A., Hendy, S. C., James, A., Parry, M., Plank, M. J., & **Steyn, N.** Effective reproduction number for COVID-19 in Aotearoa New Zealand. (2020). [https://doi.org/10.1101/2020.08.10.20172320](https://doi.org/10.1101/2020.08.10.20172320)

Binny, R. N., Hendy, S. C., James, A., Lustig, A., Plank, M. J., & **Steyn, N.** Effect of Alert Level 4 on Reff: review of international COVID-19 cases. (2020). [https://doi.org/10.1101/2020.04.30.20086934](https://doi.org/10.1101/2020.04.30.20086934)

Plank, M. J., Binny, R. N., Hendy, S. C., Lustig, A., James, A., & **Steyn, N.** A stochastic model for COVID-19 spread and the effects of Alert Level 4 in Aotearoa New Zealand. (2020). [https://doi.org/10.1101/2020.04.08.20058743](https://doi.org/10.1101/2020.04.08.20058743)

James, A., Hendy, S. C., Plank, M. J., & **Steyn, N.** Suppression and mitigation strategies for control of COVID-19 in New Zealand. (2020). [https://doi.org/10.1101/2020.03.26.20044677](https://doi.org/10.1101/2020.03.26.20044677)

>>>>

**Steyn, N.**, & Parag, K. V. (2025). Robust uncertainty quantification in popular estimators of the instantaneous reproduction number. *medRxiv preprint*. [https://doi.org/10.1101/2024.10.22.24315918](https://doi.org/10.1101/2024.10.22.24315918)

**Steyn, N.**,  Parag, K. V., Thompson, R. N., & Donnelly, C. A. (2025). A primer on inference and prediction with epidemic renewal models using sequential Monte Carlo methods. *arXiv preprint*. [https://doi.org/10.48550/arXiv.2503.18875](https://doi.org/10.48550/arXiv.2503.18875)

**Steyn, N.**, Unwin, H. J. T., Ponmattam, J., et al. (2025). Regional and national estimates of children affected by all-cause and COVID-19-associated orphanhood and caregiver death in Brazil, by age and family circumstance. *medRxiv preprint*. [https://doi.org/10.1101/2025.01.31.25321479](https://doi.org/10.1101/2025.01.31.25321479)

**Steyn, N.**, Chadeau-Hyam, M., Whitaker, M., et al. (2025). Pandemic-risk-related behaviour change in England from June 2020 to March 2022: REACT-1 study among over 2 million people. *medRxiv preprint*. [https://doi.org/10.1101/2025.03.03.25323250](https://doi.org/10.1101/2025.03.03.25323250)

@@

Watson, L. M., Plank, M. J., Armstrong, B. A., Chapman, J. R., Hewitt, J., Morris, H., Orsi, A., Bunce, M., Donnelly, C. A., & **Steyn, N**. (2024). Jointly estimating epidemiological dynamics of Covid-19 from case and wastewater data in Aotearoa New Zealand. *Communications Medicine*. [https://doi.org/10.1038/s43856-024-00570-3](https://doi.org/10.1038/s43856-024-00570-3)

@@

Binny, R. N., Priest, P., French, N. P., Parry, M., Lustig, A., Hendy, S. C., Maclaren, O. J., Ridings, K. M., **Steyn, N.**, Vattiato, G., & Plank, M. J. (2023). Sensitivity of Reverse Transcription Polymerase Chain Reaction Tests for Severe Acute Respiratory Syndrome Coronavirus 2 Through Time. *The Journal of Infectious Diseases*. [https://doi.org/10.1093/infdis/jiac317](https://doi.org/10.1093/infdis/jiac317)

Elliott, P., Whitaker, M., Tang, D., Eales, O., **Steyn, N.**, Bodinier, B., Wang, H., Elliott, J., Atchison, C., Ashby, D., Barclay, W., Taylor, G., Darzi, A., Cooke, G. S., Ward, H., Donnelly, C. A., Riley, S., & Chadeau-Hyam, M. (2023). Design and Implementation of a National SARS-CoV-2 Monitoring Program in England: REACT-1 Study. *American Journal of Public Health*. [https://doi.org/10.2105/AJPH.2023.307230](https://doi.org/10.2105/AJPH.2023.307230)

@@

Elliott, P., Eales, O., **Steyn, N.**, Tang, D., Bodinier, B., Wang, H., Elliott, J., Whitaker, M., Atchison, C., Diggle, P. J., Page, A. J., Trotter, A. J., Ashby, D., Barclay, W., Taylor, G., Ward, H., Darzi, A., Cooke, G. S., Donnelly, C. A., & Chadeau-Hyam, M. (2022). Twin peaks: the Omicron SARS-CoV-2 BA. 1 and BA. 2 epidemics in England. *Science*. [https://doi.org/10.1126/science.abq4411](https://doi.org/10.1126/science.abq4411)

Plank, M. J., James, A., Lustig, A., **Steyn, N.**, Binny, R. N., & Hendy, S. C. (2022). Potential reduction in transmission of COVID-19 by digital contact tracing systems: a modelling study. *Mathematical Medicine and Biology*. [https://doi.org/10.1093/imammb/dqac002](https://doi.org/10.1093/imammb/dqac002)

**Steyn, N.**, Lustig, A., Hendy, S. C., Binny, R. N., & Plank, M. J. (2022). Effect of vaccination, border testing, and quarantine requirements on the risk of COVID-19 in New Zealand: A modelling study. *Infectious Disease Modelling*. [https://doi.org/10.1016/j.idm.2021.12.006](https://doi.org/10.1016/j.idm.2021.12.006)

**Steyn, N.**, Plank, M. J., Binny, R. N., Hendy, S. C., Lustig, A., & Ridings, K. (2022). A COVID-19 vaccination model for Aotearoa New Zealand. *Scientific Reports* [https://doi.org/10.1038/s41598-022-06707-5](https://doi.org/10.1038/s41598-022-06707-5)

@@

Hendy, S., **Steyn, N.**, James, A., Plank, M. J., Hannah, K., Binny, R. N., & Lustig, A. (2021). Mathematical modelling to inform New Zealand’s COVID-19 response. *Journal of the Royal Society of New Zealand*. [https://doi.org/10.1080/03036758.2021.1876111](https://doi.org/10.1080/03036758.2021.1876111)

James, A., Plank, M. J., Hendy, S., Binny, R., Lustig, A., **Steyn, N.**, Nesdale, A., & Verrall, A. (2021). Successful contact tracing systems for COVID-19 rely on effective quarantine and isolation. *PLoS One*. [https://doi.org/10.1371/journal.pone.0252499](https://doi.org/10.1371/journal.pone.0252499)

**Steyn, N.**, Binny, R. N., Hannah, K., Hendy, S., James, A., Lustig, A., Ridings, K., Plank, M. J., & Sporle, A. (2021). Māori and Pacific People in New Zealand have higher risk of hospitalisation for COVID-19. *New Zealand Medical Journal*. [Link to PDF](https://nzmj.org.nz/media/pages/journal/vol-134-no-1538/maori-and-pacific-people-in-new-zealand-have-a-higher-risk-of-hospitalisation-for-covid-19/55602a1fe7-1696476650/maori-and-pacific-people-in-new-zealand-have-a-higher-risk-of-hospitalisation-for-covid-19-open-access.pdf)

James, A., Plank, M. J., Binny, R. N., Lustig, A., Hannah, K., Hendy, S. C., & **Steyn, N.** (2021). A structured model for COVID-19 spread: modelling age and healthcare inequities. *Mathematical Medicine and Biology*. [https://doi.org/10.1093/imammb/dqab006](https://doi.org/10.1093/imammb/dqab006)

Binny, R. N., Baker, M. G., Hendy, S. C., James, A., Lustig, A., Plank, M. J., Ridings, K. M., & **Steyn, N.** (2021). Early intervention is the key to success in COVID-19 control. *Royal Society Open Science*. [https://doi.org/10.1098/rsos.210488](https://doi.org/10.1098/rsos.210488)

**Steyn, N.**, Plank, M. J., James, A., Binny, R. N., Hendy, S. C., & Lustig, A. (2021). Managing the risk of a COVID-19 outbreak from border arrivals. *Journal of the Royal Society Interface*. [https://doi.org/10.1098/rsif.2021.0063](https://doi.org/10.1098/rsif.2021.0063)

Plank, M., James, A., & **Steyn, N.** (2021). Comment: weekly COVID-19 testing with household quarantine and contact tracing is feasible and would probably end the epidemic[^1]. *Royal Society Open Science*. [https://doi.org/10.1098/rsos.201546](https://doi.org/10.1098/rsos.201546)

@@

James, A., Plank, M. J., Hendy, S., Binny, R. N., Lustig, A., & **Steyn, N.** (2021). Model-free estimation of COVID-19 transmission dynamics from a complete outbreak. *PLOS ONE*. [https://doi.org/10.1371/journal.pone.0238800](https://doi.org/10.1371/journal.pone.0238800)

@@

**Steyn, N.**, Binny, R. N., Hannah, K., Hendy, S., James, A., Kukutai, T., Lustig, A., McLeod, M., Plank, M. J., Ridings, K., & Sporle, A. (2020). Estimated inequities in COVID-19 infection fatality rates by ethnicity for Aotearoa New Zealand. *New Zealand Medical Journal*. [Link to PDF](https://nzmj.org.nz/media/pages/journal/vol-133-no-1521/estimated-inequities-in-covid-19-infection-fatality-rates-by-ethnicity-for-aotearoa-new-zealand/9bb9847717-1696476126/estimated-inequities-in-covid-19-infection-fatality-rates-by-ethnicity-for-aotearoa-new-zealand-open-access.pdf)

@@

**Steyn, N.**, Plank, M. J., & Hendy, S. C. (2021). Modelling to support a future COVID-19 strategy for Aotearoa New Zealand. [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2017/01/modelling-to-support-a-future-covid-19-strategy-for-aotearoa-new-zealand.pdf)

**Steyn, N.**, Hendy, S., Plank, M., & Binny, R. (2021). Technical report: modelling the potential spread of COVID-19 during the August 2021 outbreak. [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2017/01/technical-report-28-august.pdf)

James, A., Binny, R. N., Hendy, S. C., Lustig, A., **Steyn, N.**, & Plank, M. J. (2020). Economic comparison of the use of Alert Levels 3 and 4 in eliminating the Auckland August outbreak: a cost-effectiveness analysis 21 October 2020. [Link to PDF](https://bpb-ap-se2.wpmucdn.com/blogs.auckland.ac.nz/dist/d/75/files/2020/11/Economic-comparison-Auckland-August-outbreak_forTPMwebpage_21Oct.pdf)

Binny, R. N., Hendy, S. C., James, A., Lustig, A., Plank, M. J., & **Steyn, N.** (2020). Probability of elimination for COVID-19 in Aotearoa New Zealand. [https://doi.org/10.1101/2020.08.10.20172361](https://doi.org/10.1101/2020.08.10.20172361)

Binny, R. N., Lustig, A., Brower, A., Hendy, S. C., James, A., Parry, M., Plank, M. J., & **Steyn, N.** (2020). Effective reproduction number for COVID-19 in Aotearoa New Zealand. [https://doi.org/10.1101/2020.08.10.20172320](https://doi.org/10.1101/2020.08.10.20172320)

Binny, R. N., Hendy, S. C., James, A., Lustig, A., Plank, M. J., & **Steyn, N.** (2020). Effect of Alert Level 4 on Reff: review of international COVID-19 cases. [https://doi.org/10.1101/2020.04.30.20086934](https://doi.org/10.1101/2020.04.30.20086934)

Plank, M. J., Binny, R. N., Hendy, S. C., Lustig, A., James, A., & **Steyn, N.** (2020). A stochastic model for COVID-19 spread and the effects of Alert Level 4 in Aotearoa New Zealand. [https://doi.org/10.1101/2020.04.08.20058743](https://doi.org/10.1101/2020.04.08.20058743)

James, A., Hendy, S. C., Plank, M. J., & **Steyn, N.** (2020). Suppression and mitigation strategies for control of COVID-19 in New Zealand. [https://doi.org/10.1101/2020.03.26.20044677](https://doi.org/10.1101/2020.03.26.20044677)
